Illuminating a Dark Lens : A Type Ia Supernova \\ Magnified by Galaxy Cluster Abell 2744

### 2015-04-17 authors.tex
incorporated AZ comments
cleaned up author ref macros and preamble, updated code   * adjusting figures to reflect the new redshift from the host galaxy
updates including spec fitting fig and text

\newcommand{\elsewhere}{St. Elsewhere University.}

@@

\newcommand{\Minnesota}{Department of Astronomy, University of Minnesota, 116 Church Street SE, Minneapolis, MN 55455, USA}

@@

\newcommand{\Oxford}{Department of Physics, University of Oxford, Clarendon Laboratory, Parks Road, Oxford, UK}

@@

% --------------------------------------------------
% \affilref{refcode}
%   Use this command inside each \altaffilmark{} instance.
%   It checks if the affiliation has already been used, and generates
%   a new affiliation reference number only if needed.
%   The user-defined refcode value will then be used by \affilreftxt
%   to set the same index number in the author list at the top and 
%   the footnoted list of institutional addresses. 
\makeatletter
\newcommand{\affilref}[1]{%
    \@ifundefined{c@#1}{%
    	\newcounter{#1}%
    	\setcounter{#1}{0}%
    }{}%
    \ifthenelse{\value{#1} > 0}{%
      \ref{#1}%
    }{%
      \refstepcounter{affilct}%
    	\label{#1}%
      \ref{#1}%
    	\addtocounter{#1}{1}%
    }%
}
\makeatother

% --------------------------------------------------
% \affilreftxt{refcode}{address} : 
%   Use this command after each author definition 
%   to generate the footnote text containing one institutional
%   address for that author. 
%   It checks if the address has already been used, and generates
%   a new affiliation footnote with text only if needed.
\makeatletter
\newcommand{\affilreftxt}[2]{%
    \@ifundefined{c@#1txt}{%
    	\newcounter{#1txt}%
    	\setcounter{#1txt}{0}%
    }{}%
    \ifthenelse{\value{#1txt} > 0}{%
    }{%
      \altaffiltext{\ref{#1}}{#2}%
    	\addtocounter{#1txt}{1}%
    }%
}
\makeatother
% --------------------------------------------------

@@

\author{Ryan Foley\altaffilmark{\affilref{IllinoisPhysics},\affilref{IllinoisAstro}}}
\affilreftxt{IllinoisPhysics}{\IllinoisPhysics}
\affilreftxt{IllinoisAstro}{\IllinoisAstro}

>>>>

@@

\newcommand{\Minnesota}{School of Physics and Astronomy, University of Minnesota, 116 Church Street SE, Minneapolis, MN 55455, USA}

@@

\newcommand{\Oxford}{Department of Physics, Oxford University, Keble Road, Oxford OX1 3RH, UK}

@@

% --------------------------------------------------
% \affilref{refcode}
%   Use this command inside each \altaffilmark{} instance.
%   It checks if the affiliation has already been used, and generates
%   a new affiliation reference number only if needed.
%   The user-defined refcode value will then be used by \affilreftxt
%   to set the same index number in the author list at the top and 
%   the footnoted list of institutional addresses. 
\makeatletter
\newcommand{\affilref}[1]{%
  \@ifundefined{c@#1}%
    {\newcounter{#1}%
     \setcounter{#1}{\theaffilct}%
     \refstepcounter{affilct}%
     \label{#1}%
     }{}%
  \ref{#1}%
 }
\makeatother

% --------------------------------------------------
% \affilreftxt{refcode}{address} : 
%   Use this command after each author definition 
%   to generate the footnote text containing one institutional
%   address for that author. 
%   It checks if the address has already been used, and generates
%   a new affiliation footnote with text only if needed.
\makeatletter
\newcommand*\affilreftxt[2]{%
  \@ifundefined{c@#1txt}
    {\newcounter{#1txt}%
     \setcounter{#1txt}{1}
     \altaffiltext{\ref{#1}}{#2}
     }{
     }
  }
\makeatother
% --------------------------------------------------

@@

\author{Ryan~J.~Foley\altaffilmark{\affilref{IllinoisPhysics},\affilref{IllinoisAstro}}}
\affilreftxt{IllinoisPhysics}{\IllinoisPhysics}
\affilreftxt{IllinoisAstro}{\IllinoisAstro}

### 2015-04-17 snTomas.tex
updates including spec fitting fig and text

The most probable host galaxy for SN \tomas\ is a faint and diffuse
galaxy immediately to the south-east of the SN location.  With
photometry of the host galaxy collected from the template images, we
fit the spectral energy distribution (SED) using the {\it BPZ} code --
a Bayesian photometric redshift estimator \citep{Benitez:2000}. 
%The best-fit SED template match is shown in Figure~\ref{fig:HostSED}.
From the {\it BPZ} analysis, we found the host to be most likely an
actively star-forming galaxy at a redshift of
$z=1.5\pm0.2$.  \change{This photo-z was subsequently supported by a
spectroscopic redshift of $z=1.3457\pm0.0001$, based on emission lines
consistent with \textcolor{red}{??? the [OII] $\lambda\lambda$ 3726-3729 \AA\ doublet.
???}}  Another nearby bright galaxy to the East of the SN has a
spectroscopic redshift of $z=1.742$.  This redshift was determined
from a spectrum taken with the G141 grism of the HST WFC3-IR camera,
collected as part of the Grism Lens Amplified Survey from Space
(GLASS, PI:Treu, PID:13459, \citealt{Treu:2015}). As we will see in
Sections~\ref{sec:Spectroscopy}
and \ref{sec:PhotometricClassification}, the SN data are \change{consistent
with $z=1.3457$,} meaning that the latter galaxy is a background object and
therefore has no impact on the SN magnification.

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{FIG/specfit}
\caption{  \label{fig:SpecFit}
Redshift and age determination from spectral template matching to the
the SN \tomas\ maximum light spectrum.  The y axis plots flux in
arbitrary units, and the x axis marks wavelength in \AA\ with the
observer-frame on the bottom and rest-frame on the top.
The \tomas\ spectrum observed with the
HST ACS G800L grism is shown in black, overlaid with model fits
derived from a library of Type Ia templates that have extended
rest-frame UV coverage.  {\it Top}: The best match is at $z=1.31$ with
age=$+3$ days, from the normal Type Ia SN 2012cg template when using a
smooth 3rd-order polynomial to warp the shape of the template
pseudo-continuum.  When the templates are not warped, an acceptable
fit can be found with a normal Type Ia at $z=1.31$ (middle) or at
$z\sim1$ (bottom), although the latter is inconsistent with the host
galaxy redshift prior and the light curve. No \CCSN\ templates can
provide a statistically acceptable fit at any redshift, within the age
constraints imposed by the light curve.}
\end{center}
\end{figure}

@@

To refine the redshift and phase constraints on \tomas, we next fit
the spectrum with a custom spectral matching program that employs a
library of Type Ia SN SEDs.  This library is similar to the Type Ia
spectral set used by SNID, but also includes more recent SNe with
well-observed spectral time series that extend to rest-frame UV
wavelengths (e.g. SN 2011fe and 2014J).  We first use an approach
similar to the SNID algorithm: warping the pseudo-continuum of each
template spectrum by dividing out a 3$^{\rm rd}$-order polynomial to
match the observed SED of SN \tomas.  \change{With the redshift fixed
at $z=1.3457$ we find the best fit is from the normal Type Ia SN
2014J, with a $\chi^2$ per degree of freedom $\nu$ of
$\chi^2/\nu=51.8/87$ shown in the top panel of
Figure~\ref{fig:SpecFit}.  Allowing the redshift as an additional free
parameter, we still find results that are consistent with the SNID
fits and $<2\sigma$ from the host galaxy redshift: $z=1.31\pm0.02$ and
a phase of $0\pm3$ rest-frame days. The best-fitting spectral template
in this case is the normal Type Ia SN 2012cg at $z=1.31$ with
$\chi^2/\nu=42.1/87$.}

Next, we repeat the fitting, but without any warping of the templates
to account for differences in the continuum shape.  In this iteration
we only allow each template SED to be scaled in flux coherently at all
wavelengths.  As shown in Figure~\ref{fig:SpecFit}, we again find that
when the \change{redshift is allowed as a free parameter the \tomas\
SED is best matched by a normal Type Ia SN (SN 2011fe) at redshift
$z=1.31$ with $\chi^2/\nu=102.9/90$.  Fixing the redshift to the
redshift $z=1.3457$ of the host, we also find the best match is from
SN 2011fe, with $\chi^2/\nu=133.7/91$. }

Without the continuum warping, an alternative
fit also arises: the fast-declining (91bg-like) Type Ia SN 2011iv at
$z=0.98\pm0.01$. Formally, this match provides a slightly better fit
to the unwarped \tomas\ spectrum, although the fit is notably poor at
$\sim8700$\AA\ where the most significant absorption feature is
found. Furthermore, a redshift $z\sim1$ is at odds with the host
galaxy redshift ($z=1.3457$), and we will see in the following section
that the photometric data is also incompatible with a Type Ia SN at
$z\sim1.0$.

Setting aside the $z\sim1$ solution, all other template matches
provide a consistent redshift constraint of $z=1.31\pm0.02$,
regardless of whether the templates are warped to match the SN \tomas\
continuum shape.  The inferred age from these fits is $0\pm3$
rest-frame days from peak brightness, which is also consistent with
the observed light curve.

@@

Financial support for this work was provided to SAR by NASA through grants
HST-HF-51312 and HST-GO-13386 from the Space Telescope
Science Institute, which is operated by Associated Universities for
Research in Astronomy, Inc., under NASA contract NAS 5-26555.

@@

>>>>

The most probable host galaxy for SN \tomas\ is a faint and diffuse
galaxy immediately to the south-east of the SN location.  With
photometry of the host galaxy collected from the template images, we
fit the spectral energy distribution (SED) using the {\it BPZ} code --
a Bayesian photometric redshift estimator \citep{Benitez:2000}. 
%The best-fit SED template match is shown in Figure~\ref{fig:HostSED}.
From the {\it BPZ} analysis, we found the host to be most likely an
actively star-forming galaxy at a redshift of
$z=1.5\pm0.2$.  \change{This photo-z was subsequently refined to a
spectroscopic redshift of $z=1.3457\pm0.0001$, based on optical
spectroscopy of the host (Mahler et al. in prep.) that shows two
significant ($>10\sigma$) emission lines at 8740.2 and 8746.7
Angstroms, consistent with the [OII] $\lambda\lambda$ 3726-3729 \AA\
doublet.}

The next nearest galaxy detected in HST imaging is \change{2.2\arcsec
northeast of the SN position}. It has a redshift of $z=1.742$
determined from a spectrum taken with the G141 grism of the HST
WFC3-IR camera, collected as part of the Grism Lens Amplified Survey
from Space (GLASS, PI:Treu, PID:13459, \citealt{Treu:2015}).  As we
will see in Sections~\ref{sec:Spectroscopy}
and \ref{sec:PhotometricClassification}, both the spectroscopic and
photometric data from the SN itself are \change{consistent with the
redshift of the fainter galaxy at $z=1.3457$, and
incompatible with $z=1.742$ from this brighter galaxy.}  This means
that the latter galaxy is a background object and therefore has no
impact on the SN magnification.

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{FIG/specfit}
\caption{  \label{fig:SpecFit}
Redshift and age determination from spectral template matching to the
the SN \tomas\ maximum light spectrum.  The y axis plots flux in
arbitrary units, and the x axis marks wavelength in \AA\ with the
observer-frame on the bottom and rest-frame on the top.  The \tomas\
spectrum observed with the HST ACS G800L grism is shown in black,
overlaid with model fits derived from a library of Type Ia templates
that have extended rest-frame UV coverage.  \changeG{The top two panels show
the best matching templates when using a smooth 3rd-order polynomial
to warp the shape of the template pseudo-continuum, with the redshift
fixed at z=1.3457 and then allowed to float as a free parameter.  The
lower three panels show matches found when the templates are not
warped, both with and without fixing the redshift.  The bottom panel
shows the best match in this set, although at $z=0.98$ it is
inconsistent with the host galaxy redshift prior and the light curve.}
}
\end{center}
\end{figure}

@@

To refine the redshift and phase constraints on \tomas, we next fit
the spectrum with a custom spectral matching program that employs a
library of Type Ia SN SEDs.  This library is similar to the Type Ia
spectral set used by SNID, but also includes more recent SNe with
well-observed spectral time series that extend to rest-frame UV
wavelengths (e.g. SN 2011fe and 2014J).  We first use an approach
similar to the SNID algorithm: warping the pseudo-continuum of each
template spectrum by dividing out a 3$^{\rm rd}$-order polynomial to
match the observed SED of SN \tomas.  \changeG{This approach will find
templates that have similar abundances and photospheric
velocities. With the redshift fixed at $z=1.3457$ we find the best fit
is from the normal Type Ia SN 2014J, with a $\chi^2$ per degree of
freedom $\nu$ of $\chi^2/\nu=121.8/89$ shown in the top panel of
Figure~\ref{fig:SpecFit}.  Allowing the redshift as an additional free
parameter, we still find results that are consistent with the SNID
fits and $<2\sigma$ from the host galaxy redshift: $z=1.31\pm0.02$ and
a phase of $0\pm3$ rest-frame days. The best-fitting spectral template
in this case is the normal Type Ia SN 2012cg at $z=1.31$ with
$\chi^2/\nu=98.5/88$ (second panel of Figure~\ref{fig:SpecFit}).}

\changeG{
Next, we repeat the fitting, but without any warping of the templates
to account for differences in the continuum shape.  In this iteration
we only allow each template SED to be scaled in flux coherently at all
wavelengths, so the fits are more sensitive to the overall color of
the SED. Fixing the redshift to $z=1.3457$, we find the best match is
from the normal Type Ia SN 2011fe, though the fit is quite poor, with
$\chi^2/\nu=313.2/91$ (third panel of Figure~\ref{fig:SpecFit}).  When
the redshift is allowed as a free parameter the \tomas\ SED is still
matched best by a SN 2011fe template, now at redshift $z=1.31$ with
$\chi^2/\nu=240.8/90$ (fourth panel of Figure~\ref{fig:SpecFit}).  }

Without the continuum warping, an alternative fit also arises: the
fast-declining (91bg-like) Type Ia SN 2011iv at
$z=0.98\pm0.01$. Formally, this match provides a slightly better fit
to the unwarped \tomas\ spectrum \changeG{($\chi^2/\nu=220.6/90$,
bottom panel of Figure~\ref{fig:SpecFit}),} although the fit is
notably poorer at $\sim8700$\AA\ where the most significant absorption
feature is found. Furthermore, a redshift $z\sim1$ is at odds with the
spectroscopic redshift of the nearest galaxy ($z=1.3457$), and we will
see in the following section that the photometric data is also
incompatible with a Type Ia SN at $z\sim1.0$.

Setting aside the $z\sim1$ solution, all other template matches
provide a consistent redshift constraint of $z=1.31\pm0.02$,
regardless of whether the templates are warped to match the SN \tomas\
continuum shape.  The inferred age from these fits is $0\pm3$
rest-frame days from peak brightness, which is also consistent with
the observed light curve.  \changeG{Taken together with the host galaxy
spectroscopic redshift, these fits suggest that SN \tomas\ is a normal
Type Ia SN at $z=1.3457$ with an SED color close to SN 2011fe, but
spectral absorption features similar to SN 2014J.}

@@

Financial support for this work was provided to SAR by NASA through grants
HST-HF-51312 and HST-GO-13386 from STScI, which is operated by Associated Universities for
Research in Astronomy, Inc. (AURA), under NASA contract NAS 5-26555.

@@

A.Z. acknowledges financial support from NASA through grant
HST-HF-51334.01-A awarded by STScI and operated by AURA.
